---
title: "p8105_hw3_tk2886"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(patchwork)
library(p8105.datasets)
library(ggridges)

## Set Global Options 
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6, 
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# QUESTION 1

Loading Data and Cleaning Data with janitor
```{r}
data("instacart")

instacart_df <- 
  instacart %>% 
  janitor::clean_names()
```

Analysis of instacart data
```{r}
dimension_instacart = dim(instacart_df)
instacart_row = nrow(instacart_df)
instacart_col = ncol(instacart_df)
key_variables = names(instacart_df)
```
The dimension of the instacart dataframe is `r dimension_instacart`. 
The dimensions represents the number of rows and the columns. 
The number of rows in the instacart dataframe are `r instacart_row`. 
The number of columns in the instacart dataframe are `r instacart_col`. 


Analyzing the Aisle Column/Variable and arranging most aisles with the most items 
```{r}
instacart_df %>%
  group_by(aisle) %>%
  summarise(n_obs = n()) %>%
  arrange(desc(n_obs))
```
After analyzing the data frame, there are **134 aisles**. 


Filtering the aisles with more than 10000 items ordered.
```{r}
instacart_df %>%
  group_by(aisle) %>%
  summarise(aisle_n = n()) %>%
  filter(aisle_n > 10000) %>%
  ggplot(aes(x = reorder(aisle, aisle_n), y = aisle_n)) +
  geom_bar(stat = "identity") + 
  labs(
    title = "Number of items ordered in each Aisle",
    x = "Aisle Identification",
    y = "Quantitiy purchased from each Aisle",
    caption = "Data from the Instacart package"
  ) + 
  coord_flip()
```


Creating A table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”.
```{r results='asis'}
instacart_df %>%
  filter(aisle %in%  c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle, product_name) %>%
  summarise(Product_Quantity = n()) %>%
  filter(min_rank(desc(Product_Quantity)) < 4) %>%
  arrange(aisle, desc(Product_Quantity)) %>%
  knitr::kable()
```


Showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).
```{r}
instacart_df %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarise(mean_hour = mean(order_hour_of_day)) %>%
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour
  )
```


# QUESTION 2

Loading the Data and Cleaning the Data with janitor
```{r}
data("brfss_smart2010")

behavorial_df <- 
  brfss_smart2010 %>% 
  janitor::clean_names() %>%
  filter(topic == "Overall Health") %>%
  filter(response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")) %>% 
  mutate(response = as.factor(response)) %>%
  mutate(response = forcats::fct_relevel(response, c("Poor", "Fair", "Good", "Very good", "Excellent")))
```

Analyzing 2002 for which states observed at 7 or more locations
```{r}
behavorial_df %>%
  filter(year == 2002) %>%
  group_by(locationabbr) %>%
  summarise(location_count = n_distinct(locationdesc)) %>%
  filter(location_count >= 7) %>%
  arrange(location_count)
```


Analyzing 2010 for which states observed at 7 or more locations
```{r}
behavorial_df %>%
  filter(year == 2010) %>%
  group_by(locationabbr) %>%
  summarise(location_count = n_distinct(locationdesc)) %>%
  filter(location_count >= 7) %>%
  arrange(location_count)
```


Creating a Spaghetti Plot of the average value over time within a state
```{r}
behavorial_df %>%
  filter(response == "Excellent") %>%
  group_by(year, locationabbr) %>%
  summarise(mean_data_value = mean(data_value, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = mean_data_value, color = locationabbr)) +
  geom_line(aes(group = locationabbr)) +
  labs(
    title = "Average Value throughout the year ",
    x = "Year",
    y = "Average Value",
    caption = "Data from the BRFSS"
  ) + 
  viridis::scale_color_viridis(name = "Location",
                               discrete = TRUE) +
  theme(legend.position = "right")
```


Making a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State
```{r}
behavorial_df %>%
  filter(year %in% c("2006", "2010"),
         locationabbr %in% c("NY")) %>%
  ggplot(aes(x = data_value, y = response)) + 
  geom_density_ridges(scale = .85) +
  facet_grid(~ year)
```





#Question 3
```{r}
accel_dfa <- 
  read_csv("./data/accel_data.csv") %>%
  janitor::clean_names() 

accel_dfa
```





```{r}
new_accel_df <-
  accel_dfa %>%
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minute",
    names_prefix = "activity_",
    values_to = "activity_count") %>%
  mutate(
    week = as.integer(week), 
    day_id = as.integer(day_id), 
    minute = as.integer(minute),
    hour = as.integer(minute %/% 60), 
    day = factor(day, levels = c("Sunday","Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")),
    weekday_vs_weekend = 
      case_when(
           day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday") ~ "Weekday", 
           day %in% c("Saturday", "Sunday") ~ "Weekend")) 
```



```{r}
original_accel_df = accel_dfa %>% dim()
original_accel_row = accel_dfa %>% nrow()
original_accel_col = accel_dfa %>% ncol()

key_var = new_accel_df %>% names()
accel_dim = new_accel_df %>% dim()
accel_row = new_accel_df %>% nrow()
accel_col = new_accel_df %>% ncol()
```
Both the original accelerometer data frame that was provided and the cleaned and tidy version that is created fro that specific data frame represents data for one person and this data measures “activity counts” in a short period; one-minute intervals. There is not 50,400.


Part 2
```{r}
new_accel_df %>%
  group_by(day) %>%
  summarize(total_activity = sum(activity_count, na.rm = TRUE)) %>%
  knitr::kable()
```


Part 3
```{r}
new_accel_df %>%
  group_by(day, minute) %>%
  ggplot(aes(x = minute, y = activity_count, color = day)) +
  geom_point() +
  labs(
    title = "Analysis of 24-hour Activity Per Day",
    x = "Hour in a Day (0 Hour - 24 hour)",
    y = "Total Activity Counts for Each Hour",
    caption = "Data From: Advanced Cardiac Care Center of Columbia University Medical Center") + 
  scale_x_continuous(
    breaks = c(0, 240, 480, 720, 960, 1200, 1440),
    labels = c("0hr", "4hr", "8hr", "12hr", "16hr", "20hr", "24hr"))
```

