---
title: "p8105_hw3_tk2886"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(patchwork)
library(p8105.datasets)

## Set Global Options 
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6, 
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

Loading Data and Cleaning Data with janitor
```{r}
data("instacart")

instacart_df <- 
  instacart %>% 
  janitor::clean_names()
```

Analysis of instacart data
```{r}
dimension_instacart = dim(instacart_df)
dimension_instacart
instacart_row = nrow(instacart_df)
instacart_row
instacart_col = ncol(instacart_df)
instacart_col
key_variables = names(instacart_df)
key_variables
```
The dimension of the instacart dataframe is `r dimension_instacart`. The dimensions represents the number of rows and the columns. The number of rows in the instacart dataframe are `r instacart_row`. The number of columns in the instacart dataframe are `r instacart_col`. 


Analyzing the Aisle Column/Variable and arranging most aisles with the most 
```{r}
aisle_analysis <- instacart_df %>%
  group_by(aisle) %>%
  summarise(n_obs = n()) %>%
  arrange(desc(n_obs))
```
After analyzing the data frame, there are **134 aisles**. 


Filtering the aisles with more than 10000 items ordered.
```{r}
aisle_filter <- 
  instacart_df %>%
  group_by(aisle) %>%
  summarise(aisle_n = n()) %>%
  filter(aisle_n > 10000) %>%
  ggplot(aes(x = reorder(aisle, aisle_n), y = aisle_n)) +
  geom_bar(stat = "identity") + 
  labs(
    title = "Number of items ordered in each Aisle",
    x = "Aisle Identification",
    y = "Quantitiy purchased from each Aisle",
    caption = "Data from the Instacart package"
  ) + 
  coord_flip()
```


Creating A table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”.
```{r}
inst_table1 <- instacart_df %>%
  filter(aisle %in%  c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle, product_name) %>%
  summarise(Product_Quantity = n()) 
  
  
  

view(inst_table1)
```



