---
title: "p8105_hw3_tk2886"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(patchwork)
library(p8105.datasets)

## Set Global Options 
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6, 
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# QUESTION 1

Loading Data and Cleaning Data with janitor
```{r}
data("instacart")

instacart_df <- 
  instacart %>% 
  janitor::clean_names()
```

Analysis of instacart data
```{r}
dimension_instacart = dim(instacart_df)
instacart_row = nrow(instacart_df)
instacart_col = ncol(instacart_df)
key_variables = names(instacart_df)
```
The dimension of the instacart dataframe is `r dimension_instacart`. The dimensions represents the number of rows and the columns. The number of rows in the instacart dataframe are `r instacart_row`. The number of columns in the instacart dataframe are `r instacart_col`. 


Analyzing the Aisle Column/Variable and arranging most aisles with the most 
```{r}
instacart_df %>%
  group_by(aisle) %>%
  summarise(n_obs = n()) %>%
  arrange(desc(n_obs))
```
After analyzing the data frame, there are **134 aisles**. 


Filtering the aisles with more than 10000 items ordered.
```{r}
instacart_df %>%
  group_by(aisle) %>%
  summarise(aisle_n = n()) %>%
  filter(aisle_n > 10000) %>%
  ggplot(aes(x = reorder(aisle, aisle_n), y = aisle_n)) +
  geom_bar(stat = "identity") + 
  labs(
    title = "Number of items ordered in each Aisle",
    x = "Aisle Identification",
    y = "Quantitiy purchased from each Aisle",
    caption = "Data from the Instacart package"
  ) + 
  coord_flip()
```


Creating A table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”.
```{r results='asis'}
instacart_df %>%
  filter(aisle %in%  c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle, product_name) %>%
  summarise(Product_Quantity = n()) %>%
  filter(min_rank(desc(Product_Quantity)) < 4) %>%
  arrange(aisle, desc(Product_Quantity))
```


Showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).
```{r}
instacart_df %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarise(mean_hour = mean(order_hour_of_day)) %>%
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour
  )
```


# QUESTION 2

Loading the Data and Cleaning the Data with janitor
```{r}
data("brfss_smart2010")

behavorial_df <- 
  brfss_smart2010 %>% 
  janitor::clean_names() %>%
  filter(topic == "Overall Health") %>%
  mutate(response = forcats::fct_relevel(response, c("Poor", "Fair", "Good", "Very good", "Excellent")))
```

Analyzing 2002 for which states observed at 7 or more locations
```{r}
behavorial_df %>%
  filter(year == 2002) %>%
  group_by(locationabbr) %>%
  summarise(location_count = n_distinct(locationdesc)) %>%
  filter(location_count >= 7) %>%
  arrange(desc(location_count))
```


Analyzing 2010 for which states observed at 7 or more locations
```{r}
behavorial_df %>%
  filter(year == 2010) %>%
  group_by(locationabbr) %>%
  summarise(location_count = n_distinct(locationdesc)) %>%
  filter(location_count >= 7) %>%
  arrange(desc(location_count))
```


Creating a Spaghetti Plot of the average value over time within a state
```{r}
behavorial_df %>%
  filter(response == "Excellent") %>%
  group_by(year, locationabbr) %>%
  summarise(mean_data_value = mean(data_value, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = mean_data_value, color = locationabbr)) +
  geom_line(aes(group = locationabbr)) +
  labs(
    title = "Average Value throughout the year ",
    x = "Year",
    y = "Average Value",
    caption = "Data from the BRFSS"
  ) + 
  viridis::scale_color_viridis(name = "Location",
                               discrete = TRUE) +
  theme(legend.position = "right")
```











```{r}
behavorial_df %>%
  filter(
    year %in% c("2006", "2010"), 
    locationabbr %in% c("NY")
  ) %>%
  ggplot(aes(x = locationdesc, y = data_value, color = response)) +
  geom_line(aes(group = locationdesc), alpha = 0.8) +
  facet_grid(~year) + 
  coord_flip()


  
```


